### 2025-11-21 paper/main.tex
move text upfron to methods section and discuss in more detail how number of parameters grows

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{float}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage[mode=buildnew]{standalone}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{acronym}
\usepackage{xcolor}

\input{figures/tikz-preamble.tex}

\title{Methods for Analyzing Models of\\ Rod-Shaped Bacteria}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
    \href{https://orcid.org/0009-0001-0613-7978}{
        \includegraphics[scale=0.06]{figures/orcid.pdf}
        \hspace{1mm}Jonas Pleyer
    }
    \thanks{
        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
        \href{https://cellular-raza.com}{cellular-raza.com}
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg\\
	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
	%% examples of more authors
	\And
	\href{https://orcid.org/0000-0002-6371-4495}{
        \includegraphics[scale=0.06]{figures/orcid.pdf}
        \hspace{1mm}Christian Fleck
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg
}

\newacro{abm}[ABM]{Agent-Based Model}
\newacro{ode}[ODE]{Ordinary Differential Equation}
\newacro{pbr}[PBR]{Physically-Based Rendering}
\newacro{ib}[IB]{Individual-based}
\newacro{ecoli}[\textit{E. coli}]{\textit{Escherichia coli}}
\newacro{bsubtilis}[\textit{B. subtilis}]{\textit{Bacillus subtilis}}

% Define new units
\DeclareSIUnit\pixel{pix}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
\renewcommand{\headeright}{Preprint}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}

\usepackage{enumitem}
\setlist{nolistsep}

\begin{document}
\maketitle

%###################################################################################################
\begin{abstract}
    \aclp{abm} have become popular tools in the study of complex cellular systems.
    However, despite their popularity, it is often difficult to systematically estimate their
    parameters and perform advanced analyses, e.g., using the profile-likelihood.
    In this work, we study the mechanical properties of rod-shaped bacteria and provide techniques
    for estimating their parameters and performing identifiability analysis on them.
    Our approach may serve as a blueprint for a modular framework which can be adapted to other
    research questions.
    We use time series of microscopic images as input data and incorporate advanced effects as the
    elasticity of the rod.
    The developed model serves as a generalized foundation to describe multiple species of
    rod-shaped bacteria such as \acl{ecoli} or \acl{bsubtilis}.
\end{abstract}

\textbf{Discussed Keywords (Working Title)}
\begin{itemize}
    \item Parameter Estimation
    \begin{itemize}
        \item Mechanical Properties
        \item Rod-shaped?
        \item Single-cell
    \end{itemize}
    \item \ac{abm}
    \item Methods
\end{itemize}
% keywords can be removed
% \keywords{}

\pagebreak
\renewcommand{\contentsname}{Table of Contents (remove before submission)}
\tableofcontents
\vfill
\pagebreak

% \begin{itemize}
%     \item Adapt introduction; \acp{abm} not focus of paper, but rather tool to do the job
%     \begin{itemize}
%         \item model division, growth of bacteria
%         \item \acp{abm} are good for this tool
%     \end{itemize}
%     \item Exchange $\sigma \rightarrow R_1+R_2$
%     \item Equation 8 make double norm $||\vec{F}||$
%     \item Remove auxiliary variable $\vec{F}$ only use $\vec{\nabla}V$
%     \item Table 2 in supplement? (probably not)
%     \item numerical implementation verschieben?
%     \item numerical implementation: waht do we need to simulate this?
%     \begin{itemize}
%         \item \acp{abm} are common tools to realize this
%         \item cellular\_raza is THE tool to do it
%     \end{itemize}
%     \item Methodology: single-cell parameter estimation as motivation for techniques
%     \item Which parameters are global which are individual? Discuss this in general; not in detail
%     \item depends on data, number of parameters (computational efficiency)
%     \item page 12; yes this is Discussion
%     \item page 12: maybe extend the paragraph wrt comparison in intervals
% \end{itemize}

%###################################################################################################
\section{Introduction}

% \begin{itemize}
%     \item Problem 1: How to model flexibel rods and validate these models?
%     \begin{itemize}
%         \item Flexible rods so far only covered in soft-matter studies; rarely applied in direct
%             biological contexts
%         \item \acp{abm} are \textit{the} tool to use for individual-based modeling (combine with
%             reactions, cell cycle, division, differentiation, etc.)
%     \end{itemize}
%     \item Problem 2: How to estimate parameters for individual-based models of cells/bacteria
%     \item Problem 3: How to incorporate division and heterogeneity in parameter estimation?
%
%     \item Literature Gap 1: No \acp{abm} support flexible elongated bacteria
%     \item Literature Gap 2: Noone really does parameter estimations
%     \item Literature Gap 3: Cellular heterogeneity sometimes accounted for in simulation but not
%         in parameter estimation
%
%     \item Why matter 1: No explainability without parameters
%     \item Why matter 2: How can we be sure to actually model biological reality?
%     \item Why matter 3: Bacteria especially elongated are fundamental. Understanding them is
%         important.
%
%     \item What we do 1: Construct new mechanical model for rod-shaped bacteria
%     \item What we do 2: Develop methods to compare numerical results with data
%     \item What we do 3: Apply these methods in 3 scenarios covering different aspects
%
%     \item Why contribution significant 1: Show that parameter estimation can be done
%     \item Why contribution significant 2: Show how to treat heterogeneity and that we can map 1:1
%         of model to cells
%     \item Why contribution significant 3: New model for flexible rods which integrates in
%         \texttt{cellular\_raza} and can thus be readily reused by researchers
% \end{itemize}

Bacteria come in various shapes~\cite{Young2006,Young2007}, ranging from simple spheres to complex
spirals and helices.
Elongated bacteria form an important subclass that includes some of the most studied model
organisms such as \ac{ecoli} or \ac{bsubtilis}.
Their filamentous structure underlies distinct mechanical~\cite{Daniel2003} and physiological
constraints~\cite{Jones2001}, thus influencing other cellular aspects such as
growth~\cite{Chang2014} and division~\cite{Li2007}.
It also facilitates processes such as swarming~\cite{Harshey1994} or biofilm
formation~\cite{Duvernoy2018}.

Computational methods can allow for a systematic analysis of the underlying biological reality.
To this end, \acp{abm} provide a natural way of mapping cellular processes to computational
routines.
By treating cells as individual entities with their own state, these tools enable researchers to
think in terms of cellular properties and processes, thereby allowing for a mechanistic
understanding of the system.
This bottom-up perspective has transformed the analysis of complex cellular
systems~\cite{Pleyer2023} over a wide range of
topics~\cite{Ghaffarizadeh2018,Cooper2020,Karolak2021,DeRybel2014}.
\acp{abm} are generally capable of incorporating gene regulatory networks, cell cycle, division,
differentiation, transport of extracellular compounds and many other processes which makes them the
ideal tool to study the individual behavior of cells.

Within our approach we develop a new computational model for flexible, elongated bacteria which
is a mechanical setup that most \acp{abm} are not capable of describing.
So far, only a few computational studies have considered flexible rods~\cite{Duman2018,Joshi2019},
but their models cannot be transferred directly to more complex biological systems which might
involve intracellular reactions, extracellular compounds, growth and division.
Only a few \acp{abm} support even the less complex cylindrical
shapes~\cite{Kang2014,breitwieser_biodynamo_2022,Gorochowski2012} but none are able to capture
flexible filamentous bacteria.
This poses a problem in the computational study of these systems.

Although \acp{abm} seem destined to be applied to the modeling of individual elongated bacteria,
this potential has not been fully realized yet.
Another major problem is how to perform parameter estimations for \acp{abm}.
Classical approaches in systems biology—such as maximum
likelihood~\cite{Kreutz2013,Raue2014} or maximum a posteriori estimation, identifiability analysis
and model reduction~\cite{Simpson2025,Gbor2015,Banga2008,Ashyraliyev2009}.
While some attempts have been made to properly calibrate the \ac{abm} at hand and estimate its
parameters~\cite{An2016,Lima2021,Dancik2010,Thiele2014} these methods are not widely applied and are
generally not used to explicitly capture any possible heterogeneity within the cellular population.
However, this can leave important questions unanswered:
Which combination of parameters drive the observed phenomena?
Can these parameters be identified given a set of datapoints?
Is there a simpler, i.e. reduced model that can describe the same phenomena?
If \acp{abm} are not able to answer these questions, their results will significantly lack in
explanatory power and ability to link to the specific problem at hand.

The aforementioned problems present significant gaps in the scientific literature.
First, no model can be explained or interpreted without knowing the roles and values of its
parameters or how they are constrained.
Furthermore, models which are not properly calibrated cannot guarantee to accurately represent the
underlying biological reality beyond numerical coincidence.
Finally, elongated bacteria are fundamental organisms which can be seen as a precursor to
more complex shapes such as spiral bacteria.
Obtaining a mechanistic understanding of their mechanical properties is essential to understand
emergent phenomena such as the formation of biofilms~\cite{Li2025} or bacterial
swarming~\cite{Starru2007}.

In this work, we address these challenges and aim to further close the previously identified gaps.
We construct a novel mechanical model for flexible, elongated bacteria that captures growth,
division, elastic deformations of the rod and mechanical interactions between cells.
We develop computational methods which utilize segmented microscopic images to quantitatively
compare numerical results with biological data on a single-cell level.
With these methods we can estimate the parameters of our model in three scenarios, each probing
different aspects of our model with increasing levels of complexity.

Our approach demonstrates that parameter estimation of \acp{abm} is indeed feasible, even in
the case of complex models such as single-cell modeling of flexible elongated bacteria and when
incorporating cellular heterogeneity.
The developed methods fill a methodological gap and present a novel way of fitting individual-based
models to data which can be readily adapted to various other scenarios.
We present a novel mechanical model for flexible elongated bacteria which seamlessly integrates into
the \texttt{cellular\_raza} framework, enabling researchers to extend and adapt it in future
studies.
In total, these contributions push for a more rigorous, interpretable and biolgically grounded
development process for \acp{abm} which provide deeper mechanistic insights into the nature of
cellular behaviors.

%###################################################################################################
\section{Overview}
This work consists of many subcomponents which can be combined in order to achieve various tasks.
Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
We assume that bacterial rods can be represented by a colleciton of vertices (see
section~\ref{subsec:spatial-representation}).
This assumption alone is enough to enable a cascade of other components.
Using the spatial representation, we provide visualization methods that employ a 3D rendering
engine and can produce cell masks or rendered images from a given configuration.
The data extraction component determines the vertices from given segmented microscopic images.
This allows us in particular to turn any sequence of images (movie) into a sequence of bacterial
positions and is a key feature of the parameter estimation methods.
The dynamics of the vertices are determined by a computational model which is specific
to the respective use-case and should be defined by the researcher.
It merely requires the aforementioned assumption of representing bacterial rods as a
collection of vertices.
This work will provide one example of how such a model can be defined.
Notably, the visualization and data extraction components do not require any knowledge about the
computational model, thus allowing us to reuse their functionality across differing models.

By utilizing the provided data extraction methods, we can infer properties of the bacteria that aid
us in constructing a suitable computational model.
However, the construction of any mechanistic model also requires prior konwledge which can be used
to inform our analysis methods.
For example, by extracing the positions of bacteria across a time-series of microscopic images, we
can also calculate their lengths at these time-points and thus determine their rate of growth and
type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}), thus
effectively guiding our modeling steps.
As we will show later, extracted data in the form of cellular positions or segmentation masks can
also be used for comparing computational results with biological ground truth.

Visualization methods are an integral part of any modeling endevaour.
They allow researchers to obtain feedback quickly and improve iteration times when constructing new
models.
Furthermore, we utilize them in our more advanced imaging-based comparison methods (see
section~\ref{subsec:imaging-based-comparison}) where they enable us to effectively compare
two bacterial configurations while accounting for division events.
In the future, we are planning to extend the the rendering capabilities of our approach and
generate realistic microscopic images and time-series.
Together with the parameter estimation methods, they can then be used to provide large amounts of
mechanistic data for the training of cell-segmentation and cell-tracking machine-learning
algorithms.

\begin{figure}[H]
    \centering
    \includestandalone[width=0.8\textwidth]{figures/concept/overview}
    \caption{
        The project can be divided into subcomponents that interact with and depend on
        each other.
        Every component implicitly builds on the spatial representation of rod-shaped bacteria as
        vertices (see section~\ref{subsec:spatial-representation}).
        Red boxes highlight the key scientific drivers of this study while yellow boxes indicate
        auxiliary methods.
        The blue box "Data Generation" displays a extension of this approach which is under active
        development.
    }
    \label{fig:flowchart-project-structure}
\end{figure}

%###################################################################################################
\section{A Computational Model for Rod-Shaped Bacteria}
\label{sec:computational-model}

The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
In this work, we aim to establish a model which is general enough such that it can be reused as the
basis for many variants of rod-shaped bacteria but also explicit enough such that we can quantify
its parameters.
We thus utilize only a minimal set of assumptions.
When describing the bacterial system, we distinguish between intra-cellular processes and cell-cell
interactions.
We restrict ourselves to the features discussed below and summarize them in
Table~\ref{table:simulation-aspects}.

While some rod-shaped bacteria grow by inserting new material at the ends, \ac{ecoli} and
\ac{bsubtilis} utilize a method where new material is inserted along the cylindrical part of
the rod. The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
studied as a subject on their own~\cite{Harry2001}.
On a higher level view, the rod is split apart at a point close to the center of the cell, thus
creating two new daughter cells.
It was shown that the best predictor for when the division event occurs, is the length of the rod as
opposed to the age of bacterium~\cite{Robert2014}.
Furthermore, not only the growth of the whole colony but also the growth of individual bacteria is
exponential~\cite{Amir2014,Takeuchi2005}.
The growth rates of individual bacteria and other process parameters can be heterogeneous which
should be accounted for as well \cite{Koutsoumanis2013}.

Many bacteria organize into biofilms~\cite{Dunne2002,Ong1999} which play a key role in the survival
of the collective.
Their formation is only possible due to the attractive nature of bacterial interactions between
each other~\cite{Berne2018}.
This interaction is characterized by a long-range attraction and short-ranged
adherance~\cite{Verwey1947,Trejo2013}.
We therefore need account for this phenomenon as well.

\begin{table}[H]
    \centering
    \def\arraystretch{1.3}
    \begin{tabularx}{\textwidth}{c l X}
        &\textbf{Aspect} & \textbf{Description}\\
        \toprule
        &\textbf{(C) Cellular}\\
        \midrule
        (1) & Rod-Shaped Mechanics &
            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
        (2) & Growth &
            Cells grow exponentially by inserting new material either along the circular part of the
            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
        (3) & Division &
            Rod-Shaped bacteria divide in the middle of the rod into two new agents.\\
            % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
        &\textbf{(CC) Cell-Cell Interactions}\\
        \midrule
        (4) & Adhesion &
            Bacteria attract at longer distances and adhere to each other at shorter
            distances~\cite{Verwey1947,Trejo2013}.\\
        \bottomrule
    \end{tabularx}
    \caption{
        This table lists aspects of cellular behaviour which we consider in our computational model.
        It is split into cellular aspects (C) which concern only an individual bacterium and
        interactions between cells (CC).
        These aspects are generalized in order to fit to as many types of rod-shaped bacteria as
        possible.
        These are the assumptions for the basic model that we are presenting here.
        In order for our methods to work, we are really only required to satisfy the assumption of
        our spatial representation (see section~\ref{subsec:spatial-representation}) and thus, these
        assumptions can also be loosened if required.
    }
    \label{table:simulation-aspects}
\end{table}

% --------------------------------------------------------------------------------------------------
\subsection{Spatial Representation}
\label{subsec:spatial-representation}
This work makes use of a one powerful assumption concerning the mathematical representation of the
cells.
To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
collection of auxiliary vertices $\{\vec{x}_i\}$.
We treat this collection of vertices as a discretization of the rod where the individual segments
between the vertices carry no explicit physical or biological meaning.
This assumption alone is enough to be able to build visualization and data-extraction pipelines (see
\ref{fig:flowchart-project-structure}).
Our computational model builds on top of this assumption and utilizes this representation to
define the dynamics of the vertices $\{\vec{x}_i\}$.
For this study, we will choose a particular type of model incorporating a chosen set effects.
But researchers are in general free to adapt the model for the specific problem at hand.
By building on our discretization scheme, future modeling approaches will be able to directly reuse
our established theoretical methods and computational functionalities.

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/mechanics.pdf}
        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/interaction.pdf}
        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/mie-potential-shapes.pdf}
        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \caption{
        (A) Example of an elonged bacterial rod represented by a collection of vertices which are
        connected by springs.
        The angles $\alpha_i$ will be used to model the rigidity of the rod.
        A bacterial agent which is not interacting with any external objects will tend to the
        equilibrium state of a straight rod with no curvature.
        (B) Potential curves of the Mie potential for various combinations of exponents $n,m$ with
        upper bound $\beta=2.5U_0$ and cutoff $\zeta=3(R_1+R_2)$.
    }
    \label{fig:model-bacterium}
\end{figure}

% --------------------------------------------------------------------------------------------------
\subsection{Mechanics}
\label{subsection:mechanical-model-mechanics}
In rod-shaped bacteria, the MreB protein plays a key role for the maintenance of the cellular shape.
It forms an actin-like cytoskeleton~\cite{Erickson2001,Dersch2020} and localizes specifically at
regions of negative curvature thus introducing heterogeneous cell growth~\cite{Ursell2014} which
results in a bending force during growth.
This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
external environment~\cite{Wang2010}.

We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
stiffness of the bacterial rod~\cite{Amir2014_2}.
As explained above, MReB will counteract the bending of the rod, resulting in a stiffening force.
We model the bacterium as a bending beam.
% In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988},
% but for sake of simplicity we chose to model the bacterium as a bending beam.
Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l$.
% which can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
% springs of lengths $l$.
We use the notation $\vec{w}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
between two vertices if given by

\begin{equation}
    \vec{F}_{i,\text{springs}} =
        -\gamma\left(1 - \frac{l}{\left|\vec{w}_i\right|}\right) \vec{w}_i
        + \gamma\left(1 - \frac{l}{\left|\vec{w}_{i+1}\right|}\right) \vec{w}_{i+1}.
\end{equation}

The value of the spring constants determines the elasticity with respect to elongation of the cell.
In situations where this aspect is not relevant, they merely serve as a auxiliary parameters to
maintain the length of the rod.
In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
vertex between edges is subject to a force indiced by curvature.
% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
Given the angle $\alpha_i = \sphericalangle(\vec{w}_{i-1},\vec{w}_i)$ between adjacent edges,
the bending force is proportional to the curvature $\kappa_i$ at each vertex
$\kappa_i = 2\tan(\alpha_i/2)$.
The resulting force acts along the angle bisector which can be calculated from the edge vectors.
Figure~\ref{fig:model-bacterium} (A) shows an elongated rod represented by $6$ vertices.
The force acting on vertex $\vec{x}_i$ is given by

\begin{equation}
    \vec{F}_{i,\text{curvature}} = \eta\kappa_i
        \frac{\vec{w}_i - \vec{w}_{i+1}}{|\vec{w}_i-\vec{w}_{i+1}|}
\end{equation}

where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:model-bacterium}).
The adjoining vertices $\vec{x}_{i-1},\vec{x}_{i+1}$ are subject to half of the same force in the
opposite direction $\vec{F}_{i\pm 1,\text{curvature}} = -0.5\vec{F}_{i,\text{curvature}}$.
We can see that the curvature force does not move the overall center of the rod in space.
The total force is the sum of interal and external forces.

\begin{equation}
    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}}
        + \vec{F}_{i,\text{external}}
\end{equation}

and are integrated via damped newtonian dynamics

\begin{equation}
    \partial_t^2 \vec{x}_i = \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
\end{equation}

where $\lambda$ a damping constant.
The equations have been rescaled by dividing with the mass, thus effectively eliminating this
parameter and simplifying the model.
All involved parameters and their units can be viewed in Table~\ref{table:list-of-parameters}.

% --------------------------------------------------------------------------------------------------
\subsection{Physical Interactions - Mie Potential}

Calculating the force between cells requires performing a double integral over the two polygonal
paths of the cells.
To reduce the computational costs we use an approximative, simplified model.
Given a vertex $\vec{x}_i$ on one cell, we determine the closest point $\vec{p}$ on the polygonal
line given by the vertices $\{\vec{y}_j\}$ of the interacting cell.
Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{y}_j + q\vec{y}_{j+1}$
for some specific $j$.
This approach is especially justified for short-ranged interactions.
The force is then calculated between the points $\vec{x}_i$ and $\vec{p}$ and acts on
the vertex $\vec{y}_j,\vec{y}_{j+1}$ with relative strength $(1-q)$ and $q$.

\begin{equation}
    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
\end{equation}

We utilize the Mie potential~\cite{Mie1903} to describe the attraction between two cells.
It is the generalization of the well-known Lennard-Jones~\cite{Jones1924} potential and was designed
to describe the interactions between molecules.
The Mie potential is given by:

\begin{align}
    V(r) &= V_0C\left[ \left(\frac{\sigma}{r}\right)^n -
        \left(\frac{\sigma}{r}\right)^m\right]\\
    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}
\end{align}

The strength of the potential is given by $V_0$ and the thickness of our rods
is encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the thicknesses (radii) of the
two interacting rod-like bacteria.
We utilize the shorthand notation $r=|\vec{x}_i - \vec{p}|$ for the distance between the two points
in question.
We numerically regularize the potential by imposing an upper bound $\beta$ on the norm of the
resulting force $\vec{F} = -\vec{\nabla} V$.
Furthermore, we impose a cutoff parameter $\zeta$ which sets sets the force to zero when the
distance $r$ between the interacting points exceeds this threshold.
The resulting force is given by

\begin{equation}
    \vec{F}_{i,\text{external}} = \left\{
    \begin{array}{ll}
        \beta\vec{F}/|\vec{F}| & |\vec{F}| >= \beta\\
        \vec{0} & r > \zeta\\
        \vec{F} & \text{else}
    \end{array}\right.
    % \vec{F} \theta(r-\zeta) \min\left(1, \beta / |\vec{F}|\right)
\end{equation}

where the first case ensures that the norm is bound and the second case ensures the cutoff.
A collection of configurations are displayed in figure~\ref{fig:model-bacterium} (B).

The physical interaction between the bacteria plays a crucial role for the overall dynamics of the
system.
By varying its parameters, the Mie potential is capable of describing a wide range of interaction
types from softer ones that permit some level of overlap to rigid ones with steep ascends for
distance valuels that correspond to overlaps.
In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
customizable due to the lower amount of parameters and more restrictive mathematical formulation.
Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
low amount of parameters that can be estimated.

% --------------------------------------------------------------------------------------------------
\subsection{Cell-Cycle}

To simulate cellular growth, we introduce an exponential growth term for the spring lengths $l$
\cite{Takeuchi2005}.

\begin{equation}
    \partial_t l = \mu l
\end{equation}

This process will increase the length of the cell indefinitely unless we impose a condition for the
division event.
The cell divides when the spring length $l$ has reached a fixed length threshold~\cite{Robert2014}
$l_\text{thresh}$.
To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
internal parameters in the process.
The following actions need to be taken for the old and new agent.
We first calculate the new spring lengths of the two resulting rods via

\begin{equation}
    \tilde{l} = l\left(\frac{1}{2} - \frac{R}{n_v l}\right)
\end{equation}

where $n_v$ is the number of vertices and $R$ the previously introduced thickness of the rod.
This relation ensures that that the tips of the two new rods are touching right after the division
event.
We subdivide the polygon given by the vertices of the mother cell in $n$ segments of length
$\tilde{l}$, one segment of length $2R$ and again $n$ segments of length $\tilde{l}$.
The first and latter groups determine the new vertices of the two daughter cells.
This approach assumes cells where the total length of the polygon is greater than its diameter
$n_vl>2R$.
We set the growth rates of the daughter cells explicitly during the division event.
This approach allows us to treat the growth rates of the daughter cells as full individual
parameters within our model.
Every other intracellular parameter is copied identically to the daughter cells.
Table~\ref{table:list-of-parameters} shows a list of all parameters used within our model.

\begin{table}[H]
    \centering
    \def\arraystretch{1.3}
    \begin{tabular}{l c c l c c}
        \textbf{Parameters} && \textbf{Units} & \textbf{Variables} && \textbf{Units}\\
        \toprule
        \multicolumn{4}{l}{\textbf{Mechanics}}\\
        \midrule
        Spring tension & $\gamma$ & $\unit{1\per\minute}$ & Position (vertices) & $\vec{x}_i$ & $\unit{\micro\metre}$\\
        Rigidity & $\eta$ & $\unit{\micro\metre\per\minute^2}$ & Velocity & $\partial_t\vec{x}_i$ & $\unit{\micro\metre\per\minute}$\\
        Damping & $\lambda$ & $\unit{\micro\metre}$ & Spring length & $l$ & $\unit{\micro\metre}$\\
        \multicolumn{4}{l}{\textbf{Interaction}}\\
        \midrule
        Thickness & $R$ & $\unit{\micro\metre}$\\
        Strength & $V_0$ & $\unit{\micro\metre^2\per\minute^2}$\\
        Exponent 1 & $n$ & $1$\\
        Exponent 2 & $m$ & $1$\\
        Bound & $\beta$ & $\unit{\micro\metre^2\per\minute^2}$\\
        Cutoff & $\zeta$ & $\unit{\micro\metre}$\\
        \multicolumn{4}{l}{\textbf{Growth \& Division}}\\
        \midrule
        Growth rate & $\mu$ &$\unit{\per\minute}$\\
        Spring length threshold & $l_\text{thresh}$ &$\unit{\micro\metre}$\\
        \bottomrule
    \end{tabular}
    \caption{
        List of all parameters and variables.
    }
    \label{table:list-of-parameters}
\end{table}

% --------------------------------------------------------------------------------------------------
\subsection{Numerical Implementation}
We implemented the described model in our recently developed \ac{abm} framework
\texttt{cellular\_raza}~\cite{Pleyer2025}.
By design \texttt{cellular\_raza} does not assume any particular cellular representation, which
makes it flexible and easy to adapt the simulation to particular needs.
Its design allows us to experiment with different model implementations and facilitates the
explorative capabilities of researchers.
It is furthermore able to seamlessly integrate into Python code-bases thus allowing us to utilize
methods from other popular software libraries such as \texttt{numpy} and \texttt{scipy}.
Furthermore, it is an \ac{abm} which is able to natively track mother-daughter relations of cells
over the course of a numerical simulation which is an integral component of our imaging-based
comparison algorithm (see section~\ref{subsec:imaging-based-comparison}).

%###################################################################################################
\section{Methodology}
These sections introduce methods which can be combined in order to estimate the parameters of our
computational model.
It has to be noted that they only depend on the spatial representation described in
section~\ref{subsec:spatial-representation}.
Consequently, these methods can be applied to a broader class of mechanical models that may exhibit
varied cellular behaviours.

The developed methods are independent of how we assign parameters within our model.
This means we can opt to incorporate intracellular variability by specifying parameters on a
per-cell basis.
Or we can employ shared parameters between all agents for processes which can be expected to behave
identical within margin of error.
Furthermore, mixed approaches where some parameters are assigned globally and others individually
can be realized as well.

For the presented model a list of parameters can be seen in table~\ref{table:list-of-parameters} but
the following calculations are equally true for any model with $N_p$ number of parameters.
Consider the case where we filter out a subset $N_g\leq N_p$ of parameters which will be treated
globally, thus with equal numerical value for every cell.
This means that $N_p - N_g$ parameters will be assigned individually for each of $N_c$ number of
cells.
The total amount of parameters in the system is then given by
\begin{equation}
    N_\text{tot} = N_g + N_c (N_p - N_g).
\end{equation}
It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
with the dimension ie. the total number of parameters $N_\text{tot}$.
Notice that the presented approach reflects our current work but can easily be adapted to other
scenarios where different strategies of parameters assignment are desired, e.g. when considering
two distinct species of cells which should be displayed by 2 parameters in total regardless of the
number of cells.

% --------------------------------------------------------------------------------------------------
\subsection{Visualization}
\label{subsection:visualization}
To visualize the spatial configuration of the results we employ pyvista, which provides high-level
wrappers around the popular VTK library~\cite{vtkBook,Sullivan2019}.
The initial phase of the visualization process entails the generation of three-dimensional meshes
for each bacterium as an intermediate representation.
We insert spheres matching the thickness of the rod $R$ at each vertex $\vec{x}_i$ and connect
adjacent vertices $\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, also of radius $r$ which is positioned
such that its core is identical to the connecting line segment between the two vertices.
Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.

A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
Since we use a 3D rendering engine for our visualization, the resulting image can have
overlaps of bacteria which are placed at higher or lower values along the z-direction.
To generate masks, we assign unique colours to every mesh of the respective cellular agent, ensuring
no two agents will use identical colours. \textbf{TODO semantic segmentation Citation}
We then use a parallel projection and remove any existing light sources in order to obtain a solid
color for every cell-agent.
The resulting image is shown in subfigure (B).

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/imaging-mesh.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/11571737453049821261/raw_pv/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/11571737453049821261/masks/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \caption{
        (A) Cells are rendered by combining spheres and cylinders into a single mesh.
        Spheres are highlighted in blue and cylinders in gray.
        (B) shows the result from combining sphere and cylinder meshes to obtain the shape of a
        bacterial rod.
        This render contains lighting coming from 2 light sources which thus produces a small glow
        on the upper part of the meshes.
        (C) Unique color values are assigned to each agent and light sources are removed.
        Afterwards, the image is rendered by projecting along the z-axis.
        Modifiers such as lighting or anti-aliasing are deactivated to ensure that each cell is made
        up of exactly one RGB color.
    }
    \label{fig:progression-image-generation}
\end{figure}

% --------------------------------------------------------------------------------------------------
\subsection{Position Extraction}
\label{subsec:position-extraction-algorithm}

This section will explain how we can determine the collection of vertices $\vec{x}_i$ introduced in
subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
(Figure~\ref{fig:position-extraction-algorithm} (A)).
The extracted vertices for multiple time points will be used by our parameter estimation methods as
input data.
The algorithms which we devised requires the input of cell-masks which utilize a unique color for
every cell in the image.

Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
by hand (Figure~\ref{fig:position-extraction-algorithm} (B)).
Afterwards, we proceed with the following 5 steps.
\begin{enumerate}
    \item Obtain submasks for each cell individually.
    \item Skeletonize submask with algorithm from Lee~\cite{Lee1994}
    \item Determine endpoints $\vec{q}_0,\vec{q}_1$.
        Assert that there are only 2 endpoints.
    \item Sort points such that they form a polygon.
    \item Interpolate vertices $\vec{x}_i$ from obtained polygon.
\end{enumerate}

In the first step, the image will be split into a collection of subimages that only contain
information for one single cell.
From there, we apply a skeletonization algorithm~\cite{Lee1994}
(Figure~\ref{fig:position-extraction-algorithm} (C)).
It has to be noted that there are multiple variants of skeletonization and in general, these classes
of algorithms can produce results which have intersection points and multiple endpoints, which is
undesirable in our case.
We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} works well.\\

The skeletonization algorithm supplies us with a collection of unordered lattice points $\vec{p}_i$
(pixel coordinates) where the skeleton lives.
We identify the endpoints $\vec{q}_0,\vec{q}_1$ by counting the number of neighbors in the von
Neumann neighborhood for each lattice point.
Any point with exactly one neighboring point must be and endpoint.
If we encounter more or less than two points which fit this criterion, the routine can not commence
and is aborted.
Next, we start with one of the endpoints and determine the next neighbor which was not yet picked.
We list the points in this order, thus forming a sorted polygon of pixel coordinates.\\
In the last step, we take the sorted, extracted points $\vec{p}_i$ and interpolate $n$ values for
the desired number of vertices $\vec{x}_i$.

% --------------------------------------------------------------------------------------------------
\paragraph{Benchmarking the Extraction Algorithm}

In order to determine if our extraction algorithm provides reliable results, we set out to compare
extracted data with known real data.
Our methodology makes use of the fact that we have already set up our model and solving scheme and
are able to produce cell masks from these numerical results.
In doing so, we have exact knowledge about the position of our agents since we can simply obtain
this information from our computational model.
But on the other hand we can also try to extract these positions from the generated cell-masks.
By comparing the differences between them, we get a good estimate of how well our extraction
algorithm can determine the true position of the bacteria.

In figure~\ref{fig:benchmarking-extraction-algorithm} (A-C) we see snapshots of a time series of
masks.
This particular simulation includes division events between images (B) and (C).
For all images (A-C), we have drawn the exact position of the agents with white lines and indicated
the extracted positions with crosses.
They align very well visually.
It may be hard to spot with the bare eye, but the agreement between extracted and exact positions is
visually worse at the endpoints of the rods.
To quantify this disparity, we calculate the distance between these individual vertices.
Figure~\ref{fig:benchmarking-extraction-algorithm} (D) shows the estimated length of the rods from
the extracted and exact vertices which is surrounded by the average vertex distance.
Division events lead to steps in our rod-length function since the bacterial length is halved.
We can clearly see that the agreement is very good in the beginning but right after every division
event, the uncertainty becomes large.
To illustrate this, we can see that in figure~\ref{fig:benchmarking-extraction-algorithm} (C) some
of the rods are partially overlapping which hides parts of the individual cell-masks and thus
changes the extraction result.
We also plotted the distribution of individual vertex distances in
figure~\ref{fig:benchmarking-extraction-algorithm} (E).
Darker areas in the plot correspond to distances at time-points earlier in the simulatin time.
We can see that brighter areas tend to shift towards the right, thus increasing the uncertainty of
the extraction.
However, the overall distribution shows small values of distances for a large number of vertices and
is not dominated by outliers.

\textbf{
    Die Verständlichkeit und die Formulierungen hier bedürfen noch der Verbesserung.
    Damit ich aber schneller durchkomme, verschieben wir das auf später.
}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.36\textwidth]
        {docs/source/_static/fitting-methods/algorithm/image001042.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.36\textwidth]
        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.255\textwidth]
        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        (A) Input image.
        (B) Result of the skeletonization algorithm for a single cell.
        (C) Vertices $\vec{x}_i$ which have been interpolated from the skeletonized cell.
        They represent the position of our cell-agents.
        Subfigure (D) displayes the average rod length of the exact and extracted data.
        The light blue region is the mean distance between vertices of extracted and exact
        positions which captures the inherent uncertainty of our method.
        Subfigure (E) displays all of these distances on a logarithmic scale.
        The colorscale represents earlier (darker) and later (brighter) temporal states of our
        simulation.
    }
    \label{fig:benchmarking-extraction-algorithm}
    \label{fig:position-extraction-algorithm}
\end{figure}

% --------------------------------------------------------------------------------------------------
\subsection{Parameter Estimation Methods}
\label{section:parameter-estimation}
In order to effectively estimate the parameters of our model, we need to set up a cost function
which can be minimized.
In doing so, we hope to identify all parameters, given relevant experimental data.
Parameters may be defined as global ones, which are identical across all cells or individual per
cell.
We utilize the preceding algorithms to extract data from cell-masks, simulate from an initial state
and then compare both results.
The cell-masks are easily obtainable by either using existing segmentation
tools~\cite{Cutler2022,Stringer2020,Hardo2022} or manually annotating microscopic image series.

\textbf{Du solltest hier erwähnen, dass du die Parameter per Bakterium fittest und nicht globale Parameter.}
\textit{Tue ich aber nicht grundsätzlich. Es gibt globale und individuelle parameter; also ein "mixed approach". Ich habe aber einen Hinweis eingefügt.}

\subsubsection{Direct Comparison}
\label{subsec:direct-comparison}
In the scenario where no cells divide during the observed interval, we can use our extraction
algorithm to obtain the position of the cells and then compare them directly to the positions
calculated from the computational results.
Each cell contains a collection of vertices.
Given $M$ cells with $N$ vertices each and dimension $d$, the total information can be stored in an
array $\nu_{i,j,k}$ of shape $(M,N,2)$.
Comparing the results of a simulation $\nu$ with experimental values $\tilde{\nu}$ is thus only a
matter of comparing two of these arrays within a suitable cost function~\cite{Wang2020}.
In our approach we utilize a least squares approach given by

\begin{equation}
    L_{\text{LS}}(\vec{x}, \vec{y})
        = \sqrt{\sum\limits_{i,j,k}|\nu_{i,j,k} - \tilde{\nu}_{i,j,k}|^2}
\end{equation}

which minimizes the distance between the corresponding vertices.
One important detail is that the ordering of vertices must be identical in the numerical and
experimental values.

\subsubsection{Imaging-based Comparison Algorithm}
\label{subsec:imaging-based-comparison}
The preceding algorithm is performant and effective but can not be applied in the event of a
division process.
Due to possibly differing division times, a cell which is present in the simulation might not yet
have a corresponding cell in the experimental data at this time point yet (or vice-versa) (see
Figure~\ref{fig:mask-difference-metric} (E)).
In order to still be able to compare these datapoints, we require additional methods which can be
used for comparisons across division events.
\textbf{[Es bleibt unklar, was du hier verhandeln möchtest.]}
\textit{Jetzt klarer?}

To this end we use the numerical results of the \ac{abm} simulation and our visualization
algorithms to generate cell-masks (see section~\ref{subsection:visualization}).
Each pixel has a unique color value which is directly related to one particular cell or background
(black).
We compare the generated masks to the experimentally obtained ones pixel by pixel.
Whenever two colors are not matching, a penalty of $1$ is assigned.
The total cost is the sum of all penalties.
% This algorithm is an implementation of a signed area measure.
It is important to take into account the parental relationship between the cells, by checking if
either one of the colors corresponding to the compared colors is a daughter of the other cell.
Figure~\ref{fig:mask-difference-metric} (A,B) shows two successive states of a numerical simulation
with varying amounts of cells.
We can see that some cells have divided while others have only grown without completing a division
step.
The resulting penalties without accounting for the parental relationships are displayed in
Figure~\ref{fig:mask-difference-metric} (C).
Black indicates that the pixels are matching and no penalty is assigned while a white pixel
indicates that the original pixels are not matching, thus adding to the overall cost.
In contrast, Figure~\ref{fig:mask-difference-metric} (D) takes the parental relationship into
account and highlights regions where compared pixels belong to mother-daughter cells in gray.
We assign a penalty $0\leq p \leq 1$ which is a hyperparameter for this optimization algorithm.
% It is possible to assign a custom value for the parental cost value.
Settings its value to $0$ results in a contiuous cost function in time.

Future extensions should introduce a time-dependent parent penalty $p=p(t)$ which grows with the age
of the daughter cells.
Furthermore, the developed algorithm relies on the comparison of two images and can as such make use
of existing machine-learning workflows that construct cost functions for such
scenarios~\cite{Dice1945}.\\
\textbf{@Christian are these two sentences Discussion?}

Although this optimization scheme was driven by technical challenges, it is also biologically
motivated.
In the infenitisemal time point after the division event the resulting two daughter cells will cover
precisely the space that the mother had taken up.
Often times, the exact time at which the fission event has commenced can not be determined
accurately, thus leading to uncertainties with respect to the question when the life of the daughter
cells begins and the one of the mother ends.

\textbf{Put this in Supplement and discuss it further. Still link to it here.}\\
In a different approach, one might be tempted to utilize the previously constructed direct
comparison algorithm for time intervals that do not contain any division events.
But this attempt would quickly lead to multiple problems.
It will not be able to deal with the uncertainties of the timing of the division event.
As the number of investigated cells grows, more division events will occurr even in a fixed
amount of time thus greatly decreasing or completely vanishing the effective interval at which no
division event is observed.
The latter situation is equivalent to two microscopic images as input data between which a division
event has occured.

\subsubsection{Testing the Comparison Algorithms with Division Events}
When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
data and use the generated masks to provide a benchmark of our algorithm.
To do this, we set out to calculate the cost between two successive iteration steps
$t_n$ with $t_{n+1}$.
We expect to obtain an exponential growth in the cost.
This can be explained by the fact that the growth of the total area is exponential in time.
Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
which is again an exponential function.
Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
We can see that by accounting for the parental relationship, we get the overall desired effect.
This method allows us to compare results between states that have a different number of cells.
However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
process this mask as well as the comparator, it is much more computationally demanding.
After all, for each comparison step, we need to run a full 3D visualization pipeline.
In cases where the simpler approach of comparing positions of the agents is possible, this method
should be preferred.
\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.24\textwidth]
        {docs/source/_static/fitting-methods/progressions-1.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {docs/source/_static/fitting-methods/progressions-2.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {docs/source/_static/fitting-methods/progressions-3.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {docs/source/_static/fitting-methods/progressions-4.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}\\
    \begin{minipage}{\textwidth}
        \begin{tikzonimage}[width=0.48\textwidth]
            {docs/source/_static/fitting-methods/extract_positionsdivision-comparison.png}
            \node at (0, 1)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
        \end{tikzonimage}%
        \hspace{0.01\textwidth}%
        \begin{tikzonimage}[width=0.48\textwidth]
            {docs/source/_static/fitting-methods/penalty-time-flow.png}
            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
        \end{tikzonimage}%
    \end{minipage}
    \caption{
        The figure shows two masks (A,B) which will be compared against each other.
        Image (B) contains more cells which are a result of cellular division events that have not
        yet occurred in image (A).
        (C) "Naive" comparison of cell masks, which identifies pixels of non-matching colors and
        assigns a cost by counting these pixels.
        This simple approach does not take into consideration the parental relationship.
        This approach is displayed in (D) where pixels of cells that are directly related have been
        colored in gray.
        (E) A schematic representation of the comparison problem involving multiple cells at
        different generations.
        An initial configuration produces two results by utilizing distinct methods (i.e. evoling
        our computational model in time or assigning the data for next time point).
        (F) Comparison of cost functions, exponential regression (ER) and number of cells.
        % Total cost between successive iteration steps $t_n,t_{n+1}$.
        The overall shape of the cost is expected to have the slope of an exponential function.
        We can clearly see that division events introduce jumps within this relationship.
        In the case where we account for the parental relationship, these jumps are mitigated.
        The cost value was set to $0$ for pixel-comparisons which belong to cells that are in a
        parental relationship.
    }
    \label{fig:mask-difference-metric}
\end{figure}

\section{Applications}

In this section we apply the previously developed methods to the problem of parameter estimation
of our model.
We cover 3 distinct cases with increasing level of complexity.
In the first study, we utilize results previously obtained by Amir et. al~\cite{Amir2014,Amir2014_2}
who investigated elastic deformations of \ac{ecoli} that are partially trapped within a narrow
tube and subject to a force resulting from a fluidic flow.
This setup does not involve any cell-cell interactions but specifically requires a flexible rod.
It is therefore well suited as a starting point to assess the usefulness of our methods and the
ability of our model to describe these results.\\
The last two cases consider time-series of microscopic images containing multiple cells (see
subsection~\ref{section:application-2}) which will be used to perform a parameter estimation of our
full mechanical model.
The approach is split into a first part which considers a time interval that does not contain any
division events and a second part which extends the initially chosen interval to include division
events.
In the first part, we cover a time interval which does not contain division events and thus utilize
the direct comparison algorithm while the second case contains division events and in turn requires
usage of the imaging-based comparison algorithm.

%###################################################################################################
\subsection{Application 1: Elastic deformations of \acl{ecoli}}

To assess that our model can be applied to single cells and showcase the usefulness of the
previously developed comparison techniques,
will use our previously developed computational model and apply it directly to the data provided
by the experiments of Amir et. al~\cite{Amir2014,Amir2014_2}.
The experimental setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
They studied how a single \ac{ecoli}, which is partially confined within a narrow tube, reacts
to a horizontal fluid flow that exerts a force onto the bacterium.
The experiment is particularly suited as a first attempt since it excludes cell-cell interactions
and focusses on the behaviour of a single cell.
It therefore reduces the total amount of parameters and allows for fast iteration times when
performing parameter estimations.
To model this configuration, the lowermost vertex of the rod is fixed at the base of the simulation
domain and vertices which are inside the tubulus are restricted to move only in the y-direction.

We model the force exerted by the fluid flow onto the individual vertices by
\begin{align}
    \vec{G}_i &= C
    \underbrace{|\vec{x}_i - \vec{x}_{i+1}|\sin(\beta_i)}_\text{proj. rod segment length}
    \underbrace{(yR - y^2)}_\text{flow contribution}\vec{e}_x\\
    \vec{F}_{i,\text{external}} &= \frac{1}{2}(\vec{G}_{i-1}+\vec{G}_i)
\end{align}
which is proportional to the length of the rod segment projected along the y-axis.
Similarly to Stoke's Law~\cite{Stokes1856}, it acts along the axis of the fluid flow (x-axis).
Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
The flow profile for this particular setup is proportional~\cite{Wolschin2021} to $(2yR-y^2)$ where
$y=0.5(\vec{x}_i+\vec{x}_{i+1})_y-y_b$ is the vertical displacement measured from the end of the
tube $y_b$ (see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber
along the y-axis which is hidden in Figure~\ref{fig:amir-bending-simulation} (A).
In particular, this enforces the correct boundary values the flow profile and the resulting force
$\vec{F}_{i,\text{external}}|_{y=0}=0$.
In addition we confine vertices who belong to segments that are partially within the tube
$(\vec{x}_i)_y\leq y_b$ via $(\vec{x}_{i+1})_x=x_0$ and also fix the value for the first vertex
$\vec{x}_0$ such that the bacterium is attached to the bottom.

We extract the positions of the rod from the data images (see
Figure~\ref{fig:amir-bending-simulation} (A)) and compare the simulated positions with the extracted
ones using the direct comparison algorithm (see Subsection~\ref{subsec:direct-comparison}).
Our simulation is split into two phases; One with the external flow force that initially bends the
rod and lasts $\SI{7}{\second}$ and a second phase during which the force is disabled and the cell
returns to its original shape over a span of $\SI{17}{\second}$.

Since this setup is void of any cell-cell interactions or cell division, the dynamics of the cell
can be described by a combination of growth and physical mechanics (see
subsection~\ref{subsection:mechanical-model-mechanics}).
We are thus left with 5 parameters:
Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
$\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (upright, light blue)
and intermediate (bent, dark blue) position of the rod using the optimized parameters.
The likelihood profiles displayed in figure~\ref{fig:amir-bending-simulation} (C-D, blue curves)
show that each parameter is identifiable from the given data.
However when it comes to the damping constant $\lambda$, its overall numerical value and steepness
of the curve compared to the results for the other parameters are much lower.
We can assess if the system could be described by a reduced model by setting $\lambda=0$, thus
effectively removing any damping.
The curves which are obtained in this way are displayed in orange and show that the system is also
fully identifiable in its reduced version.

It should be noted that the used images only contain snapshots of a fully-bent or fully-relaxed rod
without any intermediate steps.
This effectively limits us to effectively 2 datapoints and is likely the key factor that leads to
the previously described phenomenon.
The damping force is of dynamic nature, and takes no effect in the two steady states of the rod
being either fully bent or fully relaxed.
Gathering additional transient information would allow us to better estimate the damping parameter.

\begin{figure}[H]
    \centering
    \begin{minipage}{0.38\textwidth}
        \begin{tikzonimage}[width=\textwidth]
            {figures/amir-elastic-000032-with-axes.png}%
            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
        \end{tikzonimage}%
    \end{minipage}%
    \begin{minipage}{0.62\textwidth}
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles/drag_force.pdf}%
            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles/growth_rate.pdf}%
            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
        \end{tikzonimage}\\
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles/rod_rigidity.pdf}%
            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles/spring_tension.pdf}%
            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles/damping.pdf}%
            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
        \end{tikzonimage}%
    \end{minipage}
    \caption{
        (A) Microscopic image containing a single bacterium which is partially confined in a narrow
        tube on the bottom and free at the top where a controlled flow can be exerted on them.
        Division of the bacteria is inhibited.
        (B) A comparison of the initial and bent state show good visual agreement of our prediction
        with the experimental data.
        (C-G) Likelihood profiles for the individual parameters before and after performing model
        reduction.
    }
    \label{fig:amir-bending-simulation}
\end{figure}

%###################################################################################################
\subsection{Application 2: \acl{ecoli} growing on Agar}
\label{section:application-2}

In this section we study a collection of growing, dividing and physically interacting
\ac{ecoli}.
We use a series of microscopic images and corresponding segmentation masks for the individual
bacteria as the underlying dataset which is partially displayed in
Figure~\ref{fig:application-2-data-overview}.
With the previously developed extraction algorithm (see
section~\ref{subsec:position-extraction-algorithm}) we obtain the positions of our agents at all
time points.
These can be used as initial values and later within our parameter estimation routine.
We initialize our computational model (see section~\ref{sec:computational-model}) with the obtained
positions and optimize the parameters in order to find a set which best describes the observed
results.
The first case considers a time interval without cell-division while the second case does contain
division events.
Therefore, the first approach will optimize the direct comparison cost function (see
section~\ref{subsec:direct-comparison}) while the latter approach has to account for division events
and thus requires the imaging-based method (see section~\ref{subsec:imaging-based-comparison}).
These two scenarios present an opportunity to apply the previously developed methods within a
realistic context and validate our approach.

To show the usefulness of our previously developed methods
(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of
individually growing \ac{ecoli} using a sequence of microscopic images.
The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
section~\ref{section:data-code-availability}) has been automatically segmented with the
omnipose~\cite{Cutler2022} algorithm to obtain the required cell masks.

\textbf{Explain data more clearly. Refer to figure.}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001044.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001054.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001063.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.24\textwidth]
        {../data/crm_divide/0003/images/001074.png}%
        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \caption{TODO insert segmentation masks here}
    \label{fig:application-2-data-overview}
\end{figure}

% --------------------------------------------------------------------------------------------------
\subsubsection{The role of Heterogeneity}

We have already suspected (see Table~\ref{table:simulation-aspects}) that in order to properly
estimate parameters of our system, we need to account for natural cell-variability, thus assigning
parameters on a per-cell basis.

However, we will now present data that indicate that an individual treatment is in fact necessary.
Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
calculated from the extracted positions.
We fit a delayed exponential function to them which is defined by

\textbf{Explain fit parameter. Explain $t_0$. Give citation for exponential growth. Also link to
equation in cell-cycle section}

\begin{equation}
    l(t) =
    \left\{\begin{array}{ll}
            l_0 \exp(\lambda (t-t_0)) & t\geq t_0\\
            l_0 & \text{ else}
    \end{array}\right.
\end{equation}

where $\lambda$ is the growth rate and $t_0$ the delay.
Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
both results in good agreements (Figure~\ref{fig:estimated-growth-rates} A,B).
However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
in subfigure (C).
There, we can see much more clearly the difference between considering assigning parameters per cell
and global ones.
These results show that in order to accurately estimate the parameters of this system, we need to
account for this inherent heterogeneity and hence treat the growth parameters of the bacteria
individually.
\textbf{[Man ahnt was du meinst, doch leider ahnt man es nur. Vermeide Aussagen wie "it is clear".]}
\textit{Jetzt klarer?}

% \begin{enumerate}
%     \item motivate individual treatment of cells
%     \item pro: get at distributions of parameters; even more than growth rates
%     \item pro: probably good idea to do this in this case, since more or less required (see figure)
%     \item con: computationally more demanding
%     \item can we improve the optimization algorithm to somehow "know" that some parameters are for
%         different cells and thus do not "interact" as other parameters with each other?
%         ie. they are part of the same subspace.
% \end{enumerate}

\label{subsec:parameter-estimation-individual-treatment}
\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.28\textwidth]
        {figures/crm_fit-data-progression.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \caption{
        \textbf{TODO Fix caption!}\\
        \textbf{Link to supplement for discussion of bar plot}\\
        % TODO use same y-axis across the plots
        (B) The subfigure shows the time-evolution of the rod-lengths for every individual
        bacterium.
        These growth curves are fitted individually with delayed exponential function.
        (C) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
        fitted with an exponential function.
        The uncertainty band is the standard deviation of the averaged datapoints.
        (C) Blue dots show the combination of estimated growth rates and delays for individual
        bacteria while the orange dot corresponds to the average.
        Ellipses around the datapoints are taken from the uncertainty of the fits.
        Time is given in units of frames of the movie from which the data have been extracted.
    }
    \label{fig:estimated-growth-rates}
\end{figure}

% --------------------------------------------------------------------------------------------------
\subsubsection{Without Division}

\begin{itemize}
    \item refer to which data?
    \item refer to table of all parameters which are fitted
    \item refer to equations which are fitted
    \item "fit full mechanical model"
    \item link to everything
\end{itemize}

In this section we will focus on a time-interval which does not contain any division events.
By doing so, we can utilize the simpler and more performant direct comparison scheme (see
\ref{subsec:direct-comparison}) for our cost function.
We have chosen said interval such that effects related to the lag phase can be neglected which
allows us to sample a lower amount of paremeters thus improving runtime performance of our
optimization scheme.

% For this dataset, we assumed that all parameters which specify the interaction potential are
% identical across all agents but growth rates of the cells are individual.
We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
of the Mie potential $n,m,V_0$ a single radius $r$ for all agents and growth rates which were chosen
individually for each cell.
This leads to a total of $5+1\times 6=11$ variable parameters.
As highlighted before, in order to model non-flexible rods, we set the rigidity and spring tension
of the rod to a high, fixed value.
To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
method.
Finally, we performed a local minimization using the L-BFGS-B algorithm~\cite{Liu1989}.

Figure~\ref{fig:parameter-estimates-single-step} shows profile-likelihoods for each optimized
parameter with respect to the optimization results.
We can see that all parameters can be estimated by our optimization approach and that no
non-identifiability~\cite{Raue2009} is present.
These plots also show that the individual growth rates of the bacteria can be estimated effectively.

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.49\textwidth]
        {figures/crm_fit/snapshots/predicted-003889.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.49\textwidth]
        {figures/crm_fit/celldiffs/diff-003889.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \caption{
        (A) Final microscopic image of the time series with the predicted positions of the cells
        overlayed on top.
        (B) Differences between extracted and predicted positions.
        Note that the imaging-based comparison algorithm was not used to optimize this system but is
        chosen here as a tool to aid visualization of the prediction difference.
    }
\end{figure}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/damping.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/exponent-m.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/exponent-n.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/strength.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/radius.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}%
        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
    \end{tikzonimage}%
    \caption{
        We calculated the likelihood profiles for every optimized parameter of our model.
        All plots show how the value of the cost function varies when fixing a single parameter but
        still optimizing for the remaining ones.
        Every plot indicates that a minimum was achieved, thus every parameter could be identified.
        We show only one of the 6 individually-chosen growth rate as an example.
        The remaining plots show similar behaviour although with varying values and can be found in
        the supplementary material in section~\ref{sec:supplement-parameter-profiles}.
    }
    \label{fig:parameter-estimates-single-step}
\end{figure}

% --------------------------------------------------------------------------------------------------
\subsubsection{With Division}

In the next step, we extend our previously used time-series with successive images including
cell division events which means that we now have to use the imaging-based comparison algorithm (see
section~\ref{section:parameter-estimation}).
A practical problem in extending the time series is that the segmentation algorithm was not able to
properly segment the individual bacteria around the division event which meant that we had to omit
parts of the snapshots, leaving us with only incomplete data.

% Since we have already optimized said system, we can reduce the computational demand by choosing the
% parameters which were obtained by our previous attempts and extend them with the parameters that the
% daughter cells will carry.
We optimize for the same parameters as done before.
In addition to assigning growth rates individually, we also need to consider division lengths at
which the bacteria split and the new individual growth rates of the daughter cells as additional
parameters.
Since our simulation involves the division of 4 cells into 8 new ones, in principle, this leads
to a total of $8$ new growth rates and $4$ division lengths.
% But to decrease the size of the parameter space, we assume that the growth rates of the daughter
% cells are identical between every sibling, leaving us at only $4$ new growth rates and a total of
% $8$ parameters.
% Future approaches should aim to loosen this assumption, allowing to optimize every new growth rate.
To optimize this system, we used a similar optimization approach by again seeding with samples
generated via latin-hypercube~\cite{McKay1979}, then applying the differential evolution
algorithm~\cite{Storn1997} and finally a local minimization with Nelder-Mead~\cite{Gao2010}.
We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between
not accounting for them and assigning full weighting to this difference.
Figure~\ref{fig:celldiffs-progression-with-division} shows the difference between the cell mask as
given by the data and the cell mask predicted by the best fit of our model.
As expected, the fully-optimized system does not produce differences where the parental penalty is
required.
However, its value is frequently used within the the optimization scheme itself.

We can again calculate likelihood profiles for every optimized parameter.
The results displayed in Figure~\ref{fig:likelihood-profiles-comparison-with-division} show that
some curves exhibit a plateau close to the calculated optimum of the estimated parameter.
This concerns the division lengths and some growth rates of the mother bacteria which both determine
if and when a cell divides.
Such behaviour can be attributed to the missing images around the division events of the cells which
were not able to be segmented.
\textbf{TODO check this with results}
It can be argued that this leads to a non-identifiability for the first division length while the
remaining parameters are still identifiable although the effect is also notable.
These results highlight a reassuring finding, namely that the method of profile-likelihood is able
to properly display this lack of data.
It therefore further underlines the accuracy and predictive power of our methods.

The imaging-based comparison algorithm is computationally much more expensive compared to the
simpler position-comparison algorithm.
There are 4 main steps which are executed within any single optimization iteration that contribute
to the total wall time.
These steps are prediction via the computational model ($15.63(21)\%$), mask adjustment of the data
masks ($14.49(19)\%$), mask rendering for the numerical results ($69.2(1.4)\%$) and finally
comparison of the masks ($0.68(5)\%$) (see section \ref{section:supplement-performance}).
The second step alters the masks provided by the data for every individual numerical simulation.
This is required since we need to ensure that the colors of the data-masks match the colors of the
masks produced by our computational model.
To this end, we need to construct a family tree of the data-bacteria and compare with the
\textit{in-silico} family tree.
We can see that the generation of masks from the numerical results takes up the majority of the work.
In addition, the step of mask adjustment would not have been required within the position-comparison
algorithm.
Combining these two values, the position-comparison algorithm is $\approx 6.13(8)$ times faster.

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {figures/crm_divide/masks_diff/000000.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.32\textwidth]
        {figures/crm_divide/masks_diff/000010.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.32\textwidth]
        {figures/crm_divide/masks_diff/000020.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \caption{
        Snapshots of differences between cell masks obtained from the data and predicted by the best
        fit of our model at (A) the beginning (B) right before the division event (C) right after
        the division event.
        The optimization routine was able to fully mitigate the cost assigned due to parental
        relationships which means that the fully-optimized result contains division lengths that
        match the data.
    }
    \label{fig:celldiffs-progression-with-division}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000000.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000001.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000002.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000003.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000004.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000005.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000006.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000007.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000008.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000009.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000010.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_divide/profiles/profile-000011.png}%
        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
    \end{tikzonimage}%
    \caption{
        \text{TODO include missing profiles here}
        Likelihood profiles of every estimated parameter.
        (A-D) Division Lengths of initially placed cells.
        (E-H) Growth rates of daughter cells.
        The growth rates diplay larger values of noise within their profiles compared to the
        division lengths.
        \textbf{TODO possibly extend this description?}
    }
    \label{fig:likelihood-profiles-comparison-with-division}
\end{figure}

%###################################################################################################
\section{Discussion}

With this work we have layed out how to estimate the parameters of \acp{abm} using
microscopic image time series.
Our approach is modularized such that its structure can be applied to a variety of problems.
We introduced a \ac{ib} computational model for rod-shaped bacteria that incorporates bendiness of
the rod and is implemented with \texttt{cellular\_raza} allowing it to be flexible in design and
readily applicable for computational studies.
Furthermore, we devised novel methods to extract the spatial positions and other information of
cell-agents from masks which have been segmented from microscopic images.
The extracting method was validated using results from our computational model.
We further implemented a pipeline that can produce 3D visualizations of simulation snapshots and can
generate cell masks from data as produced by our computational model.
These imaging techniques were also utilized to design a cost function that can compare snapshots
of cellular states across cellular division events, allowing us to estimate parameters across a
large time-span.
Finally, the combination of these preceding advancements were applied to 2 case studies, which showed
that the profile-likelihood method can be used to effectively evaluate the ability of our system to
properly identify the involved parameters.
In the first example, we fit a simplified model of a singular bent rod to microscopic images.
These results highlighted that a reduction of the systems parameters is necessary in order to
identify them.
In our next step, the same system was fitted to a series of microscopic images in two secarios,
excluding and incorporating division events.
Both results showed the ability of our model to effectively estimate individual parameters of the
cells.
The case including division events produced parameter profiles with minima that were not fully able
to estimate the chosen parameters.
These results ca be traced back to a gap within the provided segmentation masks.

Our work shows that it is possible to estimate the parameters of \acp{abm} on a single-cell level.
Together with the flexible nature of \texttt{cellular\_raza}, this opens up the possibility to
design minimal \acp{abm} which can be fit to data.
The developed methods are publicly available as a well-documented python package
(\href{https://github.com/jonaspleyer/cr_mech_coli}{cr\_mech\_coli}).
The individual treatment of the cells allows us in principle to estimate the distribution of the
parameters directly without having to resort to any assumptions on the form of said distribution.
This is true for any parameter, not just the growth rates, since our approach can be generalized to
other individual cellular behaviours.

Future work should focus on multiple aspects that target challenges with respect to computational
efficiency and accuracy and extending the described model.
Additional biological effects include polar interactions~\cite{Duvernoy2018},
differentiation~\cite{vanGestel2015}, extracellular reactions~\cite{Li2025} (see also
table~\ref{table:simulation-aspects-supplement}).
When it comes to the designed algorithms, we have already discussed that the performance of
generating masks from computational results is the limiting factor for the performance of the
imaging-based comparison algorithm.
This problem could be targeted by rewriting existing code in C++ using native VTK~\cite{vtkBook}
instead of relying on the wrapper library pyvista~\cite{Sullivan2019}.
Another area of improvement is the position extraction algorithm which works well for cells living
in 2D or thin sheets but will not hold up when multiple bacteria are overlapping.
This task is well suited to be tackled by a purpose-built machine-learning tool that can either
reconstruct masks or accurately extrapolate positions from incomplete (possibly overlapping) masks.
Finally, the results of our computational model can be used to generate synthetic near-realistic
microscopic brightfield images together with their corresponding cell masks.
This combination of data allows us to train segmentation and more importantly tracking algorithms
which require large amounts of data.

In order to speed up the optimization process, it could also be valuable to initially consider all
parameters as global ones and once a certain threshold of tolerance is matched, losening this
restriction and treating cells individually.
This will allow the algorithm to initially search within a smaller space and possibly reach the
optimium more quickly in total.
Another factor that could be considered to improve performance is that individual parameters of
cells which are spatially separated will not have an immediate effect on each other, thus presenting
a loosely coupled or fully decoupled relationship.
Depening on the solver, this information could be exploited.
Ultimately, we desire to apply these methods to large amounts of data, which is done most
effectively by fitting our model to various image time-series simultanously which is yet to be
implemented.

Our approach furthermore highlighted that we are able to detect non-identifiabilities.
By choosing \texttt{cellular\_raza} as a modeling framework, we retain flexibility with respect to
model design thus effectively allowing us to reduce our model as we continue working on it.
\texttt{cellular\_raza} allows us to change the underlying representation if such unidentifiabilities
are determined to be structural (model-based) rather than practical (due to lack of
data).\textbf{TODO CITATION}
This allows us to iterate and find the model formulation which suits the data best but still retain
the mechanistic methodology of our approach.

As we continue to develop these methods, we expect to be able to reuse many of the existing
workflows\textbf{TODO better wording} such as likelihood estimation and uncertainty quantification
which have already been developed extensively for less complex systems such as \acp{ode}
\textbf{TODO CITATION}.

\begin{table}[H]
    \centering
    \begin{tabular}{l c c c c c c}
        \toprule
        % \multicolumn{6}{Growth Rates}\\
        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
        \midrule
        Fit   & 0.001405 & 0.001080 & 0.002317 & 0.001512 & 0.002048 & 0.001672\\
        \ac{abm} Direct & 0.001152 & 0.001410 & 0.001876 & 0.001683 & 0.003610 & 0.001520\\
        \ac{abm} Imaging\\
        % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
        % direct & 0.0009785162 & 0.0012641432 & 0.0021701728 & 0.0015695889 & 0.0036255661 & 0.0018419186\\
        \bottomrule
    \end{tabular}
    \caption{TODO Update and explain these values. And also add units!}
\end{table}

% \begin{itemize}
%     \item Summary:
%         "blueprint" to estimate parameters of \ac{abm},
%         bending of bacteria,
%         individual nature of cells,
%         fit model to data,
%         without+with division
%
%     \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
%         review).
%     \begin{enumerate}
%         \item polar interactions
%         \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
%         \item Proper friction; this might have a huge implication on the estimation of
%             parameters~\cite{Grant2014}.
%             This paper has already implemented this~\cite{Doumic2020}.
%             We can build on top of that.
%         \item Extracellular Reactions~\cite{Li2025}
%         \item see supplement for full list \ref{table:simulation-aspects-supplement}
%     \end{enumerate}
%     \item Data Extraction: Skeletonization Algorithm ultimately not good enough for some scenarios
%         (i.e. overlapping of cells, endpoints can sometimes be "too short")
%     \begin{itemize}
%         \item Possibly use AI for new algorithm
%     \end{itemize}
%     \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
%         from scratch with mechanistic model; train cell-segmentation and more importantly
%         cell-tracking algorithms with that.
%     \item Lack of data -> this study establishes only methods -> apply it much more broadly in
%         future
% \end{itemize}
%
% \textbf{Discussion Identifiability and Model Reduction/Flexibility}\\
%
% \textbf{Discussion Advanced Optimizations}\\
% \begin{itemize}
%     \item 
% \end{itemize}

\section{Conclusion}

This work has introduced a novel methodological framework which allows us to estimate the parameters
of models describing rod-shaped bacteria on the single-cell level by using data obtained from
microscopic images.
The approach is agnostic with respect to details of the underlying model and instead only makes use
of the fundamental spatial representation of our bacteria as a collection of vertices.
The developed functionalities are modular and can directly be reused by researchers to asses a wide
range of rod-shaped bacterial models.
Combining these methods, our workflow links experimental data to the parameter estimation of a
mechanistic, computational model and allows us to estimate the individual parameters of the
cell-agents.

Our framework allows researchers to employ well-known methods such as profile-likelihood
\textbf{TODO CITATION} and sensitivity-analysis.
The code is publicly available as the python package
\href{https://github.com/jonaspleyer/cr_mech_coli}{cr\_mech\_coli}.
The utilized cellular properties are contained within the \texttt{cellular\_raza}~\cite{Pleyer2025}
package which provides flexibility in model design, thus enabling researchers to easily reuse,
extend and combine existing building blocks.
This allows researchers to extend the model and incorporate additional biological properties.

Future work will focus on the data-generation capabilities of the software.
We will aim to generate near-realistic microscopic brightfield images together with corresponding
segmentation masks in order to train segmentation and tracking algorithms.
On the biological side, a plethora of cellular properties waits to be investigated using this novel
tecnhique.
Another open question is how to efficiently scale up this operation in order to treat large amounts
of cells such that distributions of parameters themselves can be estimated.
Put together, this work establishes a promising platform which is able to directly combine
microscopic data with single-cell agent-based modeling techniques and offers a wide range of
opportunities for future quantitative studies in cellular biolgoy.

% \begin{itemize}
%     \item Highlight methodological advancements
%     \item Constructed Model which already combines many effects; possibility to extend this even further
%     \item can now use classical methods: sensitivity-analysis, likelihood estimators and uncertainty quantification
%     \item Proof of concept: Able to fit individual parameters of the agents; can be extend this?
%     \item Provide software package which combines all functionality and is readily available
%     \item Future work:
%     \begin{itemize}
%         \item Optimize multiple data sources (images) in parallel: requires some kind of
%         "normalization" to ensure that the dimensions, pixels etc. are the same such that parameters
%         are comparable; also requires identical experimental setup
%         \item Compare different species with various modifications
%         \item Modify model futher; asymmetric forces etc.
%     \end{itemize}
% \end{itemize}

%###################################################################################################
\section{Data and Code Availability}
\label{section:data-code-availability}
All data and code which has been used in this study is available at the github repository
\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
The package which bundles the developed funtionality is published at
\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
dependency.\\
The time series of microscopic images which has been studied in section~\ref{section:application-2}
was extracted from individual frames of a video~\cite{https://doi.org/10.3203/iwf/k-129} that
showcases the exponential growth of bacteria.

\bibliographystyle{IEEEtran}
\bibliography{references}

\newpage
\appendix
%###################################################################################################
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{S\arabic{subsection}}
\section{Supplementary Material}

%###################################################################################################
\subsection{Data Preparation (TODO)}
\begin{enumerate}
    \item Have masks of data; reference image segmantation tools; or do it by hand
    \item ensure that colors are matching for bacteria between time-steps
    \item clean borders of images if needed
    \item know domain size; otherwise use pixels
\end{enumerate}

%###################################################################################################
\subsection{Parameter Estimation Results (TODO)}
\label{sec:supplement-parameter-profiles}

%TODO {explain profiles}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-5.pdf}
    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-growth-rates}
\end{figure}

%###################################################################################################
\subsection{Alternative Interaction Potential - Morse Potential}
The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
potential.
It was originally designed to describe the interaction of diatomic molecules but has since found use
in other areas~\cite{Breitwieser2021}.
It contains a repulsive and attractive part and is bound from above which makes it particularly
suitable for numerical simulations since unbound potentials can lead to large numerical deviations
thus compromising the prediction of the model.
The potential is of the form
\begin{equation}
    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
\end{equation}
and has 3 parameters.
The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
$r\geq\xi$.

%###################################################################################################
\subsection{Experimental Setup of Amir et. al}

\textbf{TODO explain some of the derivations for the flow profile etc.}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000010.png}
        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000017.png}
        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000034.png}
        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}\\
    \vspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/result-full/0000000000.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/result-full/0000000001.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/result-full/0000000002.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{F}};
    \end{tikzonimage}%
    \caption{
        (A-C) Images by Amir et. al~\cite{Amir2014_2} which have been manually segmented.
        The red line within the mask represents the extracted positions.
        (D-F) Snapshots of the numerical simulation showing the rod in its initial, fully bent and
        fully relaxed state.
    }
\end{figure}

%###################################################################################################
\subsection{How to use obtained heterogenous results}
In order to reuse this model, we also provide the option to assign growth rates by sampling from a
normal distribution with user-defined parameters.

%###################################################################################################
% \subsection{More Simulation Aspects (TODO)}
% \label{section:supplement-more-simulation-aspects}
% %TODO insert this into the discussion later
% \begin{table}[H]
%     \centering
%     \def\arraystretch{1.3}
%     \begin{tabularx}{\textwidth}{c l X}
%         &\textbf{Aspect} & \textbf{Description}\\
%         \toprule
%         &\textbf{(C) Cellular}\\
%         \midrule
%         (1) & Rod-Shaped Mechanics &
%             Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
%             approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
%         (2) & Growth &
%             Cells grow exponentially by inserting new material either along the circular part of the
%             rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
%         (3) & Differentiation &
%             \ac{bsubtilis} is known to differentiate into matrix-producing and
%             surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
%         (4) & Division &
%             The formation and continuation of van Gogh bundles is driven by cell division
%             \cite{vanGestel2015}.
%             Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
%         (5) & Variable Parameters &
%             Parameters for individual cells are not fixed values but rather taken from a
%             distribution \cite{Koutsoumanis2013}.\\
%         &\textbf{(CC) Cell-Cell Interactions}\\
%         \midrule
%         (6) & Adhesion &
%             Bacteria adhere to each other at longer distances and attach when in close contact
%             \cite{Verwey1947,Trejo2013}.
%             The interaction between rods can be polarized \cite{Duvernoy2018}.\\
%         (7) & Friction &
%             Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
%         &\textbf{(DC) Domain-Cell Interactions}\\
%         \midrule
%         (8) & External Forces &
%             Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
%             The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
%         (9) & Extracellular Reactions &
%             Bacteria can take up nutrients or possible secrete/take up signalling molecules
%             \cite{Li2025}.\\
%         \bottomrule
%     \end{tabularx}
%     \label{table:simulation-aspects-supplement}
%     \caption{
%         This table extends table~\ref{table:simulation-aspects} with additional observations.
%         It has to be noted that this table includes more specific behaviour that might not be
%         generalizable across all types of rod-shaped bacteria.
%     }
% \end{table}
% 
% \begin{itemize}
%     \item new class of aspects: (DC) Domain-Cell Interactions
%     \item List which aspects are here additionally
%     \item discuss which of them might need specific models and can not be paired with others
%     \item which can be bundled together without problems?
% \end{itemize}

%###################################################################################################
\subsection{Implementation \& Performance (TODO)}
\label{section:supplement-performance}

On the programming side, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us
to easily generate Python bindings and enable us to interoperate between the two ecosystems.

By using \texttt{cellular\_raza}, we can make use of multiple threads for running a single numerical
simulation, using the built-in parallelization methods (see Supplementary Material,
Figure~\ref{fig:performance}).

\begin{itemize}
    \item Quadratic Scaling (see figure~\ref{fig:timings-crm_divide}) comes from more dense
        simulation
    \item Overall number of agents which can be simulated is satisfactory for current purposes
    \item Improvement from $1 \Rightarrow 2$ and $2\Rightarrow4$ threads; first is major; second
        only minor
    \item Flamegraph still shows much cycles running against hurdles::Barrier::wait $\Rightarrow$
        this leaves room for improvement
    \item Flamegraph also shows that $\approx 75\%$ of all cycles is related to actual work (not
        just waiting).
        In particular to update\_mechanics functions.
\end{itemize}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.49\textwidth]
        {docs/source/_static/performance/computation-time-with-initial-agents.pdf}
    \includegraphics[width=0.49\textwidth]{docs/source/_static/performance/flamegraph.pdf}
    \caption{TODO}
    \label{fig:performance}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/crm_divide/timings.pdf}%
    \includegraphics[width=0.5\textwidth]{figures/crm_divide/time-evolution.pdf}%
    \caption{TODO}
    \label{fig:timings-crm_divide}
\end{figure}

\end{document}